Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
reintroduce orbifolds to intro, previous work

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\BH$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, we sharply bound the degree of the generators and
relations for $M(\Gamma)$, for general $\Gamma$ \todo{Peter: Why do you say general gamma?  I think you should either add "general fuchsian groups gamma" or just take out that clause, since you just defined what gamma is.}  \todo{Peter: also, its not clear to me that this should be said right here.  I think we may want to say a bit more about what's already known instead of going directly to what we have done}.  In the special that $\Gamma$ has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give similar bounds, and we extend their work to all Fuchsian groups $\Gamma$. In order to find generators and relations for $M(\Gamma)$, we translate the seemingly analytic question of understanding the ring of modular forms into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$.
Hence, for the remainder of the paper, we will work in the
algebraic category. \todo{mention that we will use the language
of Weil divisors?}

@@

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle P  \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

In particular, we would like to bound the degrees in which the
graded ring and its relations are generated.  O'Dorney
shows that the canonical ring of any $\BQ$-divisor $D =
\frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of
$\BP_k^1$ is generated in degrees up
to $e := \max(\beta_1, \beta_2)$ with relations generated in
degrees up to $2e$ \cite[Theorem 6]{dorney:canonical}. Furthermore,
O'Dorney gives a larger bound in terms of least common multiples of
the denominators when the support of $D$ is allowed to
have arbitrarily many points.
Voight and Zureick-Brown also bounded the generator and relation
degrees of log spin canonical rings of $(\sx, \Delta, \halfcan)$,
in the case that genus is positive and $\halfcan$ is effective,
by $3e$ and $6e$ for $e := \max(e_1, \ldots, e_r)$.

The goal of this paper is to provide a stricter bound on the
generator and relation degrees of log spin canonical rings
for all genera. Note that O'Dorney's bounds are asymptotically
tight for arbitrary $\BQ$-divisors on $\BP^1$, but restricting
to the log spin curve case allows for lower bounds.
The main theorem of this paper is the following improved bound on
the degree of generators and relations of log spin canonical rings:

\todo{State for even weight modular forms attributed to VZB.}
\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect field $k$, \todo{Ask David: do we need this perfectness assumption everywhere? No, use separable stacky points} so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
log spin canonical ring

\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx, \lfloor d \halfcan \rfloor)
\end{align*}

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0
$ then $R(\sx,\Delta, \halfcan)$ has relations generated in degrees
up to $\max(11, 2e_1, \ldots, 2e_r)$. Otherwise, $R(\sx, \Delta,
\halfcan)$ has relations generated in degrees up to $\max(10, 2e_1,
\ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list
of exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional}
for signatures with $g = 0$.
\end{thm}

@@

\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle P  \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\end{defn}

@@

\subsection{Junk}
\todo{add this back into the intro}
This is stuff that has been moved out of the introduction to possibly
be retooled for this section.

We can then consider the
orbifold $\Gamma \backslash \BH$ over $\BC$. For example, in the
case $\Gamma$ acts freely on $\BH$, $\Gamma \backslash \BH$ is a
Riemann Surface over $\BC$. It may well be that $\Gamma \backslash
\BH$ is not compact. Defining $\Delta$ as the divisor of cusps
associated to $\Gamma$, we can form a space $\Gamma \backslash
\BH^*$ which is a compactification of $\Gamma \backslash \BH,$ with
associated {\bf divisor of cusps} $\Delta$.

>>>>

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup of
$PSL_2(\BR)$ acting on the upper half plane $\BH$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. Often, it is useful to
bound the degree of the generators and relations for $M(\Gamma)$
for general Fuchsian groups $\Gamma$. In the special case that $\Gamma$
has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give nice bounds.
This work extends their work to all Fuchsian groups $\Gamma$.

We can now consider the orbifold $\Gamma \backslash \BH$ over $\BC$.
For example, in the case $\Gamma$ acts freely on $\BH$, $\Gamma
\backslash \BH$ is a Riemann Surface over $\BC$. It may well be
that $\Gamma \backslash \BH$ is not compact, but we can form a
space $\Gamma \backslash \BH^*$ which is a compactification of
$\Gamma \backslash \BH,$ with associated divisor of cusps $\Delta$.
In order to find generators and relations for $M(\Gamma)$, we
translate the seemingly analytic question of understanding the ring
of modular forms into the algebraic category, using a
generalization of the GAGA principle. By \cite[Proposition 6.1.5]
{vzb:stacky}, there is an equivalence of categories between orbifold
curves and log stacky curves over $\BC$. Hence, for the remainder
of the paper, we will work in the algebraic category.
\todo{mention that we will use the language of Weil divisors?}

@@

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle 
	P \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left \langle 
	\frac{1}{e_i}P_i \right \rangle \right) \subseteq \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

In particular, we would like to bound the degrees in which the
log spin canonical ring and its relations are generated. Let
$\sx$ be a stacky curve with signature $\sigma = (g; e_1, \ldots,
e_r; \delta)$. The application of O'Dorney's work to log spin
canonical rings gives a large bound for the genus zero case in
terms of least common multiples of the $e_i$'s \cite[Chapter 5]
{dorney:canonical}. Voight and Zureick-Brown bounded the 
generator degrees by $3e$ and the relation degrees by $6e$ when
$\halfcan$ is effective \cite[Corollary 10.4.6]{vzb:stacky}.

These bounds, particularly those derived from O'Dorney's work
in the genus zero case, are far from tight and do not collectively
cover all cases in all genera. The main theorem of this paper
gives significantly tighter bounds across all cases in all
genera on the generator and relation degrees of log spin canonical
rings.

\todo{State for even weight modular forms attributed to VZB.}
\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect
field $k$, so that $\sx$ has signature $\sigma = (0; e_1, \ldots,
e_r; \delta)$.

Then the log spin canonical ring is generated as a $k$-algebra by 
elements of degree at most $e = \max(5, e_1, \ldots, e_r).$ If the 
genus $g \geq 2,$ and $\Delta = 0$ then $R(\sx,\Delta, \halfcan)$
has relations generated in degrees up to $\max(11, 2e_1, \ldots,
2e_r)$. Otherwise, $R(\sx, \Delta, \halfcan)$ has relations
generated in degrees up to $\max(10, 2e_1, \ldots, 2e_r),$ so long
as $\sigma$ does not lie in a finite list of exceptional cases, as
listed in Table ~\ref{table:g-1-exceptional} for signatures with $g
= 1$ and Table ~\ref{table:g-0-exceptional} for signatures with $g =
0$.
\end{thm}

@@

\[
	\di \sx = \left(\bigoplus_{P\notin \{P_1,\ldots, P_r\}} \langle 
	P \rangle \right) \oplus \left(\bigoplus_{i = 1}^r \left \langle 
	\frac{1}{e_i}P_i \right \rangle \right) \subseteq \BQ \otimes \di X.
\]
\end{defn}

@@
